---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# stabm
[![R-CMD-check](https://github.com/bommert/stabm/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/bommert/stabm/actions/workflows/R-CMD-check.yaml)
[![CRAN Status](https://www.r-pkg.org/badges/version-ago/stabm)](https://cran.r-project.org/package=stabm)
[![DOI](https://joss.theoj.org/papers/10.21105/joss.03010/status.svg)](https://doi.org/10.21105/joss.03010)


`stabm` provides an implementation of many measures which assess the stability of feature selection.
The following stability measures are currently included:

```{r}
stabm::listStabilityMeasures()
```


## Installation

You can install the released version of stabm from [CRAN](https://cran.r-project.org/package=stabm) with:

```{r, eval=FALSE}
install.packages("stabm")
```
For the development version, use [devtools](https://cran.r-project.org/package=devtools):

```{r, eval = FALSE}
devtools::install_github("bommert/stabm")
```


## Contributions

This R package is licensed under the [LGPL-3](https://www.gnu.org/licenses/lgpl-3.0.en.html).
If you encounter problems using this software (lack of documentation, misleading or wrong documentation, unexpected behaviour, bugs, ...) or just want to suggest features,
please open an issue in the [issue tracker](https://github.com/bommert/stabm/issues).
Pull requests are welcome and will be included at the discretion of the author.


## Code of Conduct

Please note that the `stabm` project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html).
By contributing to this project, you agree to abide by its terms.


## Citation

If you use stabm, please cite our [JOSS article](https://doi.org/10.21105/joss.03010):
```{r echo = FALSE, comment = ""}
toBibtex(citation("stabm"))
```
